The Impact of Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2014-09-18 draft.tex
Second draft version to send

\begin{figure}	%fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/fig1.eps}
\caption{Optical spectra of three nova-like variables: 
RW Sex (Top; Beuermann et al. 1992),
IX Vel (Middle; Beuermann \& Thomas 1990)
and RW Tri in eclipse (Bottom; Groot et al. 2004).
The data have been digitized from the respective publications. 
These systems have inclinations of $30^\circ$, $60^\circ$ and $70^\circ$ 
respectively.
The trend of increasing Balmer line emission with inclination can be seen.
In RW Tri strong single-peaked emission in the Balmer lines is seen even
in eclipse, indicating that the lines may be formed in a spatially extensive disk wind, and there is even a suggestion
of a (potentially wind-formed) recombination continuum in the eclipsed spectrum.}
\label{novalikes}
\end{figure}

@@

Our aim here is not
to provide a `best fit' model to match any particular dataset.
Rather, we take
a successful UV wind model as a starting point
and examine its effect 
when one utilises the better treatment of lines provided
by the macro-atom scheme.

% The specific model 
% parameters are summarised in table 1, and follow LK02. 
% They are motivated by a need to agree with typical observed values 
% for NLs and the values required to reproduce the UV line profile

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_beta.eps}}   
}
\caption{
Line profiles for \ha\ (right) and \hb\ (left) with varying kinematic 
properties, computed for an inclination of $80^\circ$.
The benchmark model is shown in the top panel, and also in grey in each 
subsequent plot for comparison. The parameter that has been altered from the benchmark model 
is marked on each panel. For the left and middle right columns, models lower down
are more slowly accelerating and have denser wind bases. The middle right column
shows the variation with wider launch radii, and the right hand column shows 
how a boundary layer of increasing temperature affects the spectrum.
% How the H$\alpha$ and H$\beta$ lines in our model change with the kinematic properties
% of the wind. The three separate panels show three different values of $\dot{M}_W$, 
% while $\alpha$ and $R_v$ are all gradually increased from top to bottom and left to right respectively.
% Larger values of these quantities increase the density at the base of the wind. There is a clear transition
% from double peaked to single peaked behaviour, and at the highest densities the wind
% essentially becomes neutral as the optically thick regions self-shield the wind from photoionizing radiation.
% \bf{Each of these figures will highlight the next gen model in bold.
% and contain a grid like the above showing a pair of spectral features in
% the left and right panels. The grey shaded line will show the benchmark model
% for easy comparison.  
% The most promising parameters at the moment are acceleration exponent and launch radii.}
% }
}
\label{halpha}
\end{figure} %fullpage

@@

The left panel of figure~\ref{jump} shows the evolution
of the Balmer jump with our kinematic parameters. Higher density models
produce sufficient recombination continuum emission to fill 
in the Balmer jump. This offers a potential solution to a 
longstanding problem when reconciling theoretical
models of NL spectra with observations, as originally suggested by
\cite{KLWB98}. The fact that this emission is very sensitive to the parameters
that govern the density at the wind base also supports
the idea that a transition region close to the disk plane is necessary
to explain the emission features present \citep{kd1997}.

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_1640_0.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_4686.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \heiiuv\ (left) and \heiiopt\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{heiifig}
\end{figure}

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_balmer.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_uv_0.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{The evolution of the Balmer jump and higher order Balmer lines (left) 
and the \civfull\ line profile (right)
with the kinematic properties of the wind,
computed for an inclination of $80^\circ$. 
the panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

@@

\begin{figure} %fullpage
\includegraphics[width=0.8\textwidth]{figures/fig13_eclipse.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from Model B.}
\label{rwtricomp}
\end{figure} %fullpage

>>>>

\begin{figure}	%fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/fig1.eps}
\caption{Optical spectra of three nova-like variables: 
RW Sex (Top; Beuermann et al. 1992),
IX Vel (Middle; Beuermann \& Thomas 1990)
and RW Tri in eclipse (Bottom; Groot et al. 2004).
The data have been digitized from the respective publications and a few of the more
prominent emission and absorption lines are marked. 
These systems have inclinations of $30^\circ$, $60^\circ$ and $70^\circ$ 
respectively.
The trend of increasing Balmer line emission with inclination can be seen.
In RW Tri strong single-peaked emission in the Balmer lines is seen even
in eclipse, indicating that the lines may be formed in a spatially extensive disk wind, and there is even a suggestion
of a (potentially wind-formed) recombination continuum in the eclipsed spectrum.}
\label{novalikes}
\end{figure}

@@

Our aim here is not
to provide a `best fit' model to match any particular dataset.
Rather, we take
a successful UV wind model as a starting point
and examine its effect 
when one utilises the better treatment of lines provided
by the macro-atom scheme.
The model parameters are shown in table 1. 
they are identical to LK02, and were chosen both to reflect typical
values for NLs and reproduce realistic UV line profiles.
% The specific model 
% parameters are summarised in table 1, and follow LK02. 
% They are motivated by a need to agree with typical observed values 
% for NLs and the values required to reproduce the UV line profile

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_beta.eps}}   
}
\caption{
Line profiles for \ha\ (left) and \hb\ (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$.
The benchmark model is shown in the top left panel, and the improved optical
model described in section 6 is highlighted in bold. 
The parameter that has been altered from the benchmark model 
is marked in each column and row. 
Models towards the bottom right
are have more slowly accelerating winds and denser wind bases. 
% The middle right column
% shows the variation with wider launch radii, and the right hand column shows 
% how a boundary layer of increasing temperature affects the spectrum.
% How the H$\alpha$ and H$\beta$ lines in our model change with the kinematic properties
% of the wind. The three separate panels show three different values of $\dot{M}_W$, 
% while $\alpha$ and $R_v$ are all gradually increased from top to bottom and left to right respectively.
% Larger values of these quantities increase the density at the base of the wind. There is a clear transition
% from double peaked to single peaked behaviour, and at the highest densities the wind
% essentially becomes neutral as the optically thick regions self-shield the wind from photoionizing radiation.
% \bf{Each of these figures will highlight the next gen model in bold.
% and contain a grid like the above showing a pair of spectral features in
% the left and right panels. The grey shaded line will show the benchmark model
% for easy comparison.  
% The most promising parameters at the moment are acceleration exponent and launch radii.}
% }
}
\label{halpha}
\end{figure} %fullpage

@@

The left panel of figure~\ref{jump} shows the evolution
of the Balmer jump with our kinematic parameters. 
The `absorption' feature just redward of the Balmer jump
is artificial. It is caused by line blanketing in the stellar atmospheres,
which is not filled in by line emission from the wind
due to the use of a 20-level Hydrogen atom.
The higher density models
produce sufficient recombination continuum emission to fill 
in the Balmer jump. This offers a potential solution to a 
longstanding problem when reconciling theoretical
models of NL spectra with observations, as originally suggested by
\cite{KLWB98}. The fact that this emission is very sensitive to the parameters
that govern the density at the wind base also supports
the idea that a transition region close to the disk plane is necessary
to explain the emission features present \citep{kd1997}.
% \noindent\rule{16cm}{0.4pt}

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_1640.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_4686.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \heiiuv\ (left) and \heiiopt\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{heiifig}
\end{figure}

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_balmer.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_uv.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{The evolution of the Balmer jump and higher order Balmer lines (left) 
and the \civfull\ line profile (right)
with the kinematic properties of the wind,
computed for an inclination of $80^\circ$. 
the panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

@@

\begin{figure} %fullpage
\includegraphics[width=0.8\textwidth]{figures/fig13_eclipse.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from Model B.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure} %fullpage
